Fixing Vulnerabilities Hinders Maintainability

### 2019-02-03 paper/ICPC19/icpc19.tex
polishing ack
another issue fixed

In the example (Listing~\ref{lst:vuln}), all functions receive as input the \textbf{handshake} represented by \texttt{s}. Every SSL/TLS connection begins with a handshake which is reponsible by the negotiation between two parties. The handshake determines what cryptographic algorithms will be used to encrypt communications, verify the server and establish a secure connection before beginning data transfer. In the \texttt{dtls1\_reassemble\_fragment} function the value of
\texttt{msg\_hdr-$>$frag\_off+frag\_len} is checked against the maximum
handshake message size, but \texttt{msg\_len} is allocated to the fragment
buffer in line $37$. Thus, if the fragment was within the allowed size, the
pending handshake message could consume $18MB$ ($16MB$+$2MB$). According to the
message commit, $10$ handshake messages are allowed, so the attacker could
consume until $180MB$ per DTLS connection.

@@

The impact of each type of security refactoring on the maintainability of
open-source software is presented in Figure~\ref{fig:pat}. The results of the
Wilcoxon signed-rank test by pattern returned statistical evidence which
supports that the maintainability of open-source software decreases for 4
different patterns: Broken Autentication (p = 0.018), Denial-of-Service
(p = 0.029), Memory Leak (p = 0.009) and Miscellaneous (p = $3.07$x$10^{-4}$).
The results also returned statistical significance that evidence that maintainability
is more likely to not suffer differences for open-source software after removing
Components with Known Vulnerabilities (p = $7.96$x$10^{-4}$) and fixing Cross-Site
Scripting Vulnerabilities (p = 0.001). Injection and Cross-Site Request Forgery
produced more cases where the refactorings had a negative impact in the maintainability.
However, the results obtained for these patterns are not statistically significant.

Projects with Broken Authentication flaws are the most affected ones by security
refactorings. 55.3\% (21) of the refactorings had a negative impact in the maintainability
while 35\% (11) had a positive impact and only 13.8\% (6) did not suffer any
change. The results also show that fixing memory leaks affects the maintainability
of 51.3\% (41) of the cases and leaves only 13.8\% (11) of the cases unaffected.
In 35\% (28) of the cases, maintainability increased. Refactoring vulnerabilities
that allow denial-of-service attacks (e.g., CVE-2014-3506) has a negative impact
of 47.5\% (19) in the maintainability of open-source software. 30.0\% (12) of
the cases exhibit maintainability improvement while 22.5\% (9) show that overall
maintainability does not suffer any change. Miscellaneous refactorings pattern
integrates several types of security flaws whose impact is of 43.1\% (84) negative
in the maintainability of open-source sofware. 31.8\% (62) of the cases do not
exhibit any change in maintainability and 25.1\% (49) of cases have a positive
impact on the maintainability of open-source software.

It is also important to notice that overall the maintainability of open-source
software is not affected by Cross-Site Scripting vulnerabilities and by Using
Components with Know Vulnerabilities. Cross-Site Scripting does not suffer any
decrease in the maintainability of 55.3\% (63) of cases and 21.1\% (24) have a
positive effect on it. Only 23.7\% (27) of the cases hinder the software
maintainability. From 21 cases evaluated for Using Componentes with Known
Vulnerabilities, only 4.8\% (1) has a positive effect in the maintainability
of software. 14.3\% (3) harms maintainability software and 81\% has no effect
in the maintainability.

@@

\textbf{\textit{Security refactorings harm the maintainability of open-source software.}}

We found that 40.2\% of the security refactoring have a negative impact on the
maintainability of open-source software, Figure~\ref{fig:secvsreg}. This raises
a new tradeoff when developers need to fix vulnerabilities on their projects
because they might be affecting the software maintainability or even introducing
new vulnerabilities. This study exhibits evidence that developers may have to
reduce maintainability for the sake of safety. We understand that developers
should be able to produce secure code without affecting the maintainability of
their projects. But there are still some concerns that need to be addressed:
\begin{itemize}
	\item \textbf{Security Frameworks} need to be secure and user-friendly~\cite{5287006, 7676144}. Unfortunately,
	there are several vulnerabilities that stay undisclosed. Thus, developers should not
	trust them blindly and check defaults.
	\item \textbf{Documentation} of software should provide developers with the best practices to implement safe and maintainable software. Unmaintainable code is difficult to test, analyze and re-use.
	\item \textbf{Proper Authorization Mechanisms} should be designed and implemented. This is
	one of the most complex security features to implement. Some mechanisms might assist authorization maintenance in long run, such as the combination of \texttt{HTTP} filters with \texttt{Spring Security} annotations.
	\item \textbf{Deprecated code} should be avoided by developers. They should integrate tools such as BCH on their projects CI/CD pipelines to help them be more aware of the maintainability of their code and issues (e.g., dead and duplicated code).
	\item\textbf{Programming Languages} should provide new mechanisms to easily implement security patterns without endangering maintainability.

@@

The impact of a refactoring depends on its complexity, i.e., if the refactoring
adds complexity to the codebase it is probably affecting the software
maintainability. One evidence of that is the fact of \emph{Cross-Site
Scripting} and \emph{Using Components with Know Vulnerabilities} refactorings
endure more cases without any impact in the maintainability. Typically, these
type of vulnerabilities are one line fixes --- adding an \texttt{escape}
function or changing the version of the component, respectively. Whereas,
Broken Authentication and vulnerabilities such as the one presented in
Section~\ref{sec:motivation} responsible by DoS attacks are more difficult to
fix (several code lines deleted and added). Another example is the Apple
authentication flaw discovered in 2017 (CVE-2017-13872)\footnote{CVE-2017-13872
details available at https://support.apple.com/en-us/HT208315 (Accessed on
\today{})}, where any user could login as root with an empty password. Patrick
Wardle examined the cause of the issue\footnote{\emph{Why $<$blank$>$ Gets You
Root} available at https://objective-see.com/blog/blog\_0x24.html (Accessed on
\today{})} and conclude that the flaw was due to the introduction of high
cyclomatic complexity in a method used to verify the password. In sum,
\emph{Cross-Site Scripting} and \emph{Using Components with Known
Vulnerabilities} should take part of the developers preoccupations along with
the other patterns that explicitly hinder maintainability because despite their
low complexity these are patterns that still produce a considerable negative
impact on the maintainability. Real efforts have been made in creating and
incorporating coding standards and best practices in software production.
However, security is still far from perfection and fixing vulnerabilities might
hinder software maintainability. Thus, it is important to provide tool support
to developers to help them applying these patterns without harming software
maintainability.

@@

This study also exhibits evidence that some patterns need more attention than
others. Broken Autentication issues, Denial-of-Service attacks and Memory Leaks
endure maintainability in 55.3\%, 51.3\% and 47.5\%, respectively. Following
the implications of our work, we draw attention for the need of secure security
frameworks, code best practices documentation to produce safe and maintainable
code, proper authorization mechanisms design and implementation, better tools
to introduce in the CI/CD pipeline to help software maintenance and new or
improved programming languages to allow developers to easily implement safety
in their software without endangering maintainability.

>>>>

In the example (Listing~\ref{lst:vuln}), all functions receive as input the \textbf{handshake} represented by \texttt{s}. Every SSL/TLS connection begins with a handshake which is reponsible by the negotiation between two parties. The handshake determines what cryptographic algorithms will be used to encrypt communications, verify the server and establish a secure connection before beginning data transfer. In the \texttt{dtls1\_reassemble\_fragment} function the value of
\texttt{msg\_hdr-$>$frag\_off+frag\_len} is checked against the maximum
handshake message size, but \texttt{msg\_len} is allocated to the fragment
buffer in line $37$. Thus, if the fragment was within the allowed size, the
pending handshake message could could use up to $18MB$ ($16MB$+$2MB$). According to the
message commit, $10$ handshake messages are allowed, so the attacker could
consume until $180MB$ per DTLS connection.

@@

The impact of each type of security refactoring on the maintainability of
open-source software is presented in Figure~\ref{fig:pat}. The results of the
Wilcoxon signed-rank test by pattern returned statistical evidence which
supports that the maintainability of open-source software decreases for 4
different patterns: \emph{Broken Autentication} (p = 0.018), \emph{Denial-of-Service}
(p = 0.029), \emph{Memory Leak} (p = 0.009) and \emph{Miscellaneous} (p = $3.07$x$10^{-4}$).
The results also returned statistical significance that evidence that maintainability
is more likely to not suffer differences for open-source software after removing
\emph{Components with Known Vulnerabilities} (p = $7.96$x$10^{-4}$) and fixing \emph{Cross-Site
Scripting Vulnerabilities} (p = 0.001). \emph{Injection} and \emph{Cross-Site Request Forgery}
produced more cases where the refactorings had a negative impact in the maintainability.
However, the results obtained for these patterns are not statistically significant.

Projects with \emph{Broken Authentication} flaws are the most affected ones by security
refactorings. 55.3\% (21) of the refactorings had a negative impact in the maintainability
while 35\% (11) had a positive impact and only 13.8\% (6) did not suffer any
change. The results also show that fixing \emph{Memory Leak}s affects the maintainability
of 51.3\% (41) of the cases and leaves only 13.8\% (11) of the cases unaffected.
In 35\% (28) of the cases, maintainability increased. Refactoring vulnerabilities
that allow \emph{Denial-of-Dervice} attacks (e.g., CVE-2014-3506) has a negative impact
of 47.5\% (19) in the maintainability of open-source software. 30.0\% (12) of
the cases exhibit maintainability improvement while 22.5\% (9) show that overall
maintainability does not suffer any change. \emph{Miscellaneous} refactorings pattern
integrates several types of security flaws whose impact is of 43.1\% (84) negative
in the maintainability of open-source sofware. 31.8\% (62) of the cases do not
exhibit any change in maintainability and 25.1\% (49) of cases have a positive
impact on the maintainability of open-source software.

It is also important to notice that overall the maintainability of open-source
software is not affected by \emph{Cross-Site Scripting} vulnerabilities and by \emph{Using
Components with Know Vulnerabilities}. \emph{Cross-Site Scripting} does not suffer any
decrease in the maintainability of 55.3\% (63) of cases and 21.1\% (24) have a
positive effect on it. Only 23.7\% (27) of the cases hinder the software
maintainability. From 21 cases evaluated for \emph{Using Componentes with Known
Vulnerabilities}, only 4.8\% (1) has a positive effect in the maintainability
of software. 14.3\% (3) harms maintainability software and 81\% has no effect
in the maintainability.

@@

\textbf{\textit{Security refactorings harm the maintainability of open-source software.}}
%
We found that 40.2\% of the security refactoring have a negative impact on the
maintainability of open-source software, Figure~\ref{fig:secvsreg}. This raises
a new tradeoff when developers need to fix vulnerabilities on their projects
because they might be affecting the software maintainability or even introducing
new vulnerabilities. This study exhibits evidence that developers may have to
reduce maintainability for the sake of safety. We understand that developers
should be able to produce secure code without affecting the maintainability of
their projects. But there are still some concerns that need to be addressed:
\begin{itemize}
	\item \textbf{Security Frameworks} need to be secure and user-friendly~\cite{5287006, 7676144}. Unfortunately,
	there are several vulnerabilities that stay undisclosed. Thus, developers should not
	trust them blindly and check defaults.
	\item \textbf{Documentation} of software should provide developers with the best practices to implement safe and maintainable software. Unmaintainable code is difficult to test, analyze and re-use.
	\item \textbf{Proper Authorization Mechanisms} should be designed and implemented. This is
	one of the most complex security features to implement. Some mechanisms might assist authorization maintenance in long run, such as the combination of \texttt{HTTP} filters with \texttt{Spring Security} annotations.
	\item \textbf{Deprecated code} should be avoided by developers. They should integrate tools such as BCH on their projects CI/CD pipelines to help them be more aware of the maintainability of their code and issues (e.g., dead and duplicated code).
	\item\textbf{Programming Languages} should provide new mechanisms to easily implement security patterns without endangering maintainability.

@@

The impact of a refactoring depends on its complexity, i.e., if the refactoring
adds complexity to the codebase it is probably affecting the software
maintainability. One evidence of that is the fact of \emph{Cross-Site
Scripting} and \emph{Using Components with Know Vulnerabilities} refactorings
endure more cases without any impact in the maintainability. Typically, these
type of vulnerabilities are one line fixes --- adding an \texttt{escape}
function or changing the version of the component, respectively. Whereas,
\emph{Broken Authentication} and vulnerabilities such as the one presented in
Section~\ref{sec:motivation} responsible by DoS attacks are more difficult to
fix (several code lines deleted and added). Another example is the Apple
authentication flaw discovered in 2017 (CVE-2017-13872)\footnote{CVE-2017-13872
details available at https://support.apple.com/en-us/HT208315 (Accessed on
\today{})}, where any user could login as root with an empty password. Patrick
Wardle examined the cause of the issue\footnote{\emph{Why $<$blank$>$ Gets You
Root} available at https://objective-see.com/blog/blog\_0x24.html (Accessed on
\today{})} and conclude that the flaw was due to the introduction of high
cyclomatic complexity in a method used to verify the password. In sum,
\emph{Cross-Site Scripting} and \emph{Using Components with Known
Vulnerabilities} should take part of the developers preoccupations along with
the other patterns that explicitly hinder maintainability because despite their
low complexity these are patterns that still produce a considerable negative
impact on the maintainability. Real efforts have been made in creating and
incorporating coding standards and best practices in software production.
However, security is still far from perfection and fixing vulnerabilities might
hinder software maintainability. Thus, it is important to provide tool support
to developers to help them applying these patterns without harming software
maintainability.

@@

This study also exhibits evidence that some patterns need more attention than
others. \emph{Broken Autentication} issues, \emph{Denial-of-Service} attacks and \emph{Memory Leaks}
endure maintainability in 55.3\%, 51.3\% and 47.5\%, respectively. Following
the implications of our work, we draw attention for the need of secure security
frameworks, code best practices documentation to produce safe and maintainable
code, proper authorization mechanisms design and implementation, better tools
to introduce in the CI/CD pipeline to help software maintenance and new or
improved programming languages to allow developers to easily implement safety
in their software without endangering maintainability.
